---
title: "Numerical Solution for Interaction Equilibrium"
format: html
toc: true
toc-depth: 3
number-sections: true
lang: en
---


|   There are generally two main approaches for solving mathematical problems in the Natural Sciences: analytical solutions and numerical solutions. An *analytical solution* (theoretical model) relies on a function containing one or more independent variables, a dependent variable, and one or more parameters to be determined, which together describe the problem situation. In contrast, a *numerical solution* does not depend on an explicit descriptive equation, but rather on approximations aimed at minimizing a general function (*F(x) = 0*). Numerical methods are employed when the theoretical model is complex, when an explicit solution is not desired, or when such a solution does not exist.
\

## Equation:

|   The *site1* program numerically solves the quantities involved in a bimolecular interaction equilibrium:

```{=tex}
\begin{equation}
L + R \begin{array}{c}
_{k_{on}}\\
 \rightleftharpoons\\
^{k_{off}}
\end{array} LR 
\end{equation}
```

\

*Where:*

* *R* = free receptor (protein, nucleic acid, etc.);
* *LR* = formed complex;
* *k$_{on}$* = association rate constant of the complex;
* *k$_{off}$* = dissociation rate constant of the complex.

|   The equation describing complex formation is:

```{=tex}
\begin{equation}
K_d = \frac{[RL]}{[R] + [L]} = \frac{k_{off}}{k_{on}}
\end{equation}
```

|   Where *K$_d$* represents the *equilibrium dissociation constant* of the complex. Thus:

$$
LR = \frac{L \cdot R}{K_d}
$$

|   Accordingly, for the numerical solution of the theoretical model described above, the minimization functions are defined as:

$$
F(x)=0:\quad R_0 = R + LR ;;\rightarrow;; R + \frac{L \cdot R}{K_d} - R_0 = 0
$$

$$
F(x)=0:\quad L_0 = L + LR ;;\rightarrow;; \frac{L \cdot R}{K_d} - L_0 = 0
$$

## site1

|   The program uses the *MSLV* (*Multiple Equation Solver*) command for the simultaneous solution of multiple equations, together with the [MSLV2](https://www.hpcalc.org/hp49/math/numeric/mslv2v100.zip) library, which optimizes its performance. The library can also be loaded onto the calculator from its source code provided by the developer.
\

|   Program inputs and outputs are as follows:

```{r, eval=FALSE}
# Inputs: 
  L0, R0, and Kd

# Outputs:
  L × R plot;
  Matrix of L and LR values on the stack;
  Variables (L, R, RL, initial values, and seeds).
```

## Files:

1. [Site1 program](Site1.zip)
2. [Program source code](Site1.pdf)
   \

## Example of use:

```{r, eval=FALSE}
1. Place on the stack:
  a. L0 = 10;
  b. R0 = 5;
  c. Kd = 10;

2. Run "Site1".
```

*Note:*

1. Although fully functional, the program is relatively slow when compared to desktop software. The example above required approximately 16 s to complete graphical output and variable display, whereas the program [Octave](https://octave.org/) required only 1 s.

## References {.unnumbered}

1. Olson, John S. *Numerical analysis of kinetic ligand binding data*. Methods in Enzymology, Vol. 76. Academic Press, 1981, pp. 652–667.
